Constraining the Interiors of Asteroids Through Close Encounters

### 2022-08-19 paper/appendices.tex
Responding to Julien's comments
Revised paper

To extract the density distribution of an asteroid, the main text assumes that the angular velocity data of the asteroid is observable. It is possible that the orientation of the asteroid may be more readily available as a data set. In this appendix, we generate an orientation data set for the reference asteroid flyby, extract density moments from it, and compare the results to moments extracted from angular velocity data.

@@

Two models were discussed in section \ref{sec:density-distro} to translate density moment constraints into density distribution constraints. Here we outline two additional models --- the nearly-uniform and the harmonic models --- which are less conventional but still useable for extracting density distribution properties. Unlike the finite element and lumpy models discussed in the main text, these models will yield smooth distributions with no discrete transitions.

\subsection{Nearly-uniform model}
In this ``nearly-uniform'' model, we seek to pick one density distribution from the many distributions consistent with the data by maximizing a prior distribution $f[\rho(\bm r)]$, which can be chosen manually. Any prior distribution can be chosen, but the following prior is both interesting and numerically efficient.

@@

\section{Animated density distributions}

This appendix contains animations to better display the density distributions shown in the main text. Each frame represents a cross section perpendicular to the $\unit z$-axis, starting with negative $z$ and ending with positive $z$. All densities are divided by the mean asteroid density.

\begin{figure*}
  \textbf{Please see the published version of the paper for these animations, or find them online at the following links.}

  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/asym-fe-d.mp4}{FE asymmetric density} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/asym-fe-s.mp4}{FE asymmetric deviation} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/asym-fe-u.mp4}{FE asymmetric uncertainty} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/asym-fe-r.mp4}{FE asymmetric significance}

  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/asym-l-d.mp4}{Lumpy asymmetric density} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/asym-l-s.mp4}{Lumpy asymmetric deviation} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/asym-l-u.mp4}{Lumpy asymmetric uncertainty} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/asym-l-r.mp4}{Lumpy asymmetric significance}

  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sym-fe-d.mp4}{FE symmetric density} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sym-fe-s.mp4}{FE symmetric deviation} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sym-fe-u.mp4}{FE symmetric uncertainty} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sym-fe-r.mp4}{FE symmetric significance}

  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sym-l-d.mp4}{Lumpy symmetric density} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sym-l-s.mp4}{Lumpy symmetric deviation} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sym-l-u.mp4}{Lumpy symmetric uncertainty} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sym-l-r.mp4}{Lumpy symmetric significance}

  \caption{Density distributions extracted via the finite element model for the asymmetric (\textit{top two rows}) and symmetric (\textit{bottom two rows}) reference asteroids. The finite element model (\textit{first and third rows}) and the lumpy model (\textit{second and fourth rows}) are employed. From left to right, the densities, deviations from the true density, uncertainties, and significance of the deviations are plotted. Animated form of figure \ref{fig:den-uniform}.}
  \label{fig:animated-uniform}
\end{figure*}

\begin{figure*}
  \textbf{Please see the published version of the paper for these animations, or find them online at the following links.}

  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/move-fe-d.mp4}{FE density} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/move-fe-s.mp4}{FE deviation} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/move-fe-u.mp4}{FE uncertainty} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/move-fe-r.mp4}{FE significance}

  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/move-l-d.mp4}{Lumpy density} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/move-l-s.mp4}{Lumpy deviation} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/move-l-u.mp4}{Lumpy uncertainty} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/move-l-r.mp4}{Lumpy significance}

  \caption{Density distributions extracted via the finite-element (\textit{top}) and lumpy (\textit{bottom}) models for an asteroid with an off-center core. From left to right, the densities, deviations from the true density, uncertainties, and significance of the deviations are plotted. Animated form of figure \ref{fig:den-move}.}
  \label{fig:animated-move}
\end{figure*}

\begin{figure*}
  \textbf{Please see the published version of the paper for these animations, or find them online at the following links.}

  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sph-3-fe-d.mp4}{FE density} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sph-3-fe-s.mp4}{FE deviation} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sph-3-fe-u.mp4}{FE uncertainty} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sph-3-fe-r.mp4}{FE significance}

  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sph-3-l-d.mp4}{Lumpy density} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sph-3-l-s.mp4}{Lumpy deviation} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sph-3-l-u.mp4}{Lumpy uncertainty} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/sph-3-l-r.mp4}{Lumpy significance}

  \caption{Density distributions extracted via the finite-element (\textit{top}) and lumpy (\textit{bottom}) models for an asteroid with a centred core. From left to right, the densities, deviations from the true density, uncertainties, and significance of the deviations are plotted. Animated form of figure \ref{fig:den-sph}.}
  \label{fig:animated-sph}
\end{figure*}

\begin{figure*}
  \textbf{Please see the published version of the paper for these animations, or find them online at the following links.}

  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/double-fe-d.mp4}{FE density} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/double-fe-s.mp4}{FE deviation} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/double-fe-u.mp4}{FE uncertainty} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/double-fe-r.mp4}{FE significance}

  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/double-l-d.mp4}{Lumpy density} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/double-l-s.mp4}{Lumpy deviation} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/double-l-u.mp4}{Lumpy uncertainty} \hfill
  \href{https://github.com/jack-dinsmore/asteroid-tidal-torque/tree/main/paper/gifs/double-l-r.mp4}{Lumpy significance}

  \caption{Density distributions extracted via the finite-element (\textit{top}) and the two-lump lumpy (\textit{bottom}) models for an asteroid with two counterbalancing cores. From left to right, the densities, deviations from the true density, uncertainties, and significance of the deviations are plotted. Animated form of figure \ref{fig:den-double}.}
  \label{fig:animated-double}
\end{figure*}

>>>>

To extract the density distribution of an asteroid, the main text assumes that the angular velocity data of the asteroid is observable. It is possible that the orientation of the asteroid may be more readily available as a data set, notably if a large collection of radar antenna fail to follow the encounter sufficiently. In this appendix, we generate an orientation data set for the reference asteroid flyby, extract density moments from it, and compare the results to moments extracted from angular velocity data.

@@

Two models were discussed in section \ref{sec:density-distro} to translate density moment constraints into density distribution constraints. Here we outline two additional models which are less conventional but still useable for extracting density distribution properties. Unlike the finite element and lumpy models discussed in the main text, these models will yield smooth distributions with no discrete transitions.

\subsection{Nearly-uniform model}
In this ``nearly-uniform'' model, we pick one density distribution from the many distributions consistent with the data by maximizing a prior distribution $f[\rho(\bm r)]$. Any prior distribution can be chosen, but the following prior is both interesting and numerically efficient.

@@

### 2022-08-19 paper/main.tex
Revised paper

Over the past twenty years, the increase in quantity and quality of sensitive all-sky surveys has prompted the discovery of numerous asteroids. Such advances have been made via ground-based surveys such as the Catalina Sky Survey \cite{larson1998catalina}, Pan-STARRS \cite{kaiser2002pan}, and the Lincoln Near-Earth Asteroid Research project (LINEAR) \cite{stokes2000lincoln}, as well as space-based instruments such as the Wide-field Infrared Survey Explorer (WISE) mission \cite{wright2010wide}. Many of these asteroids are relatively small, but some are kilometre-sized and a few are predicted to closely encounter Earth or other planets in the near future. More encounter candidates are likely to be discovered by new efforts such as the Large-aperture Synoptic Survey Telescope (LSST) \cite{tyson2002large}. Their encounters can then be monitored by global ground-based networks such as the Las Cumbres Observatory (LCO) \cite{brown2013cumbres}. Such ground-based monitoring is typically used to derive the rotation period of an asteroid and its surface properties (see e.g. \cite{10.1093/mnras/stab1252}).

Since the tidal torque acting on an asteroid during an encounter depends on the interior mass distribution, the careful monitoring of angular velocity variations during an encounter also presents a window into the interior properties of asteroids. The gravitational two-body system has been studied in the context of tidal torque to different orders and with several different methods \cite{paul88, SCHEERES2000106, ashenberg07, BOUE2009750, HouMar2017}. Further theoretical studies showed that the tidal torque, observed through angular velocity perturbations, is sensitive to asteroid interior density distribution \cite{Naidu_2015, Makarov2022ChaosOO, scheeres2004evolution}.

Angular velocity perturbations have been observed and used to extract asteroid properties in several cases, including for the 2013 encounter of (367943) Duende with Earth \cite{MOSKOVITZ2020113519, benson2020spin}, and asteroid binaries (3905) Doppler and (617) Patroclus \cite{DESCAMPS2020113726, BERTHIER2020113990}. Orbital and physical properties, including moment of inertia (MOI) ratios have also been extracted for asteroid 99942 Apophis, discovered on June 19, 2004 by R. A. Tucker, D. J. Tholen, and F. Bernardi \cite{giorgini2005recent, giorgini2008predicting, smalley20052004}, and on target to encounter Earth in 2029 \cite{yu2014numerical, hirabayashi2021finite, valvano2022apophis, Lee2022Apophis}. However, density distribution features beyond the MOI ratios have not yet been extracted for any asteroid encounters. More research is needed to study in what cases these effects are observable, and what factors generally inhibit observation of these new features. It seems pivotal to augment previous work on the affect of tidal torque on Apophis' angular velocity in particular \cite{souchay2014rotational, souchay2018changes} so that upcoming observations may constrain these properties and thus improve our predictions.

We address this community need by developing a methodology to translate (1) time series of asteroid angular velocity data into constraints on density moments and (2) constraints on density moments into constraints on an asteroid's density distribution. Other techniques, such as measurement of tidal distortion \cite{RICHARDSON199847}, impact or seismometry experiments \cite{RICHARDSON2005325}, or gradiometry \cite{carroll2018tidal}, may additionally constrain the density distribution. In section \ref{sec:methods}, we introduce the analytical and numerical fundamentals of this methodology. There, we describe a simulation used to integrate the equations of motion and produce synthetic data of angular velocity over time, followed by a Markov Chain Monte Carlo (MCMC) fit process which  extracts density moments from the fit data. We then describe two methods to generate full density distributions from the density moments. In section \ref{sec:results}, we present the results of a series of injection-retrieval tests demonstrating the extent to which the properties of an asteroid chosen to generate synthetic spin data can be retrieved via our methodology. Finally, in section \ref{sec:discussion}, we assess the sensitivity of these constraints to various physical, observational, and methodological parameters to provide guidance for monitoring upcoming close encounters.

@@

\subsection{Extracting density moments from spin data}
\label{sec:fit}
Given synthetic data, an Affine Invariant MCMC Ensemble sampler was used to generate Posterior Probability Distributions (PPDs) from flat priors. We use the Python implementation \texttt{emcee} \cite{foreman2013emcee}. Our parameters were $\gamma_0$, $K_{20}$, $K_{22}$, and $K_{3m}$ (10 in total), bounded by $|\gamma_0| < \pi/4$, and the $K_{2 m}$ bounds given in equation \ref{eqn:parameter-bounds}. Computing $K_{3m}$ for many positive density distributions revealed a $|K_{3m}| \lesssim 0.01$ typically, so we use more conservative bounds on $K_{3 m}$ of $|K_{3 m}| < 1$. In general, spin data is most sensitive to $\gamma_0$ and $K_{2m}$, which we call the ``first-order parameters.'' We call $K_{3m}$ the ``second-order parameters.''

@@

Before the MCMC was run, local minima in the likelihood were found via the Nelder-Mead algorithm implemented in \texttt{scipy} \cite{Gao2012}. Generally, only one local minimum existed, except when $K_{22}=0$ in which case rotational symmetry caused all values of $\gamma_0$ to be degenerate. Ensemble walkers were initialized near the local minimum, distributed by a Gaussian approximation of the likelihood, as determined via the Hessian of the likelihood at the minimum. Due to the high sensitivity of the angular velocity data to density moments, the minimization procedure sometimes failed to isolate the minimum likelihood. Therefore, a simpler simulation without the $K_{3m}$ terms of equation \ref{eqn:tidal-torque} was first used to minimize likelihood as a function of the first-order parameters $\gamma_0$ and $K_{2m}$, and then the full simulation was used to find the second-order parameters $K_{3m}$ with the first-order parameters fixed.

@@

The translation rules of spherical harmonics \cite{Gelderen1998TheSO} give that the density moments in the hybrid frame are
\begin{equation}
  \begin{aligned}
    K_{\ell m} = &\brackets{\sum_{i=0}^N \frac{a_i^2 + r_i^2}
    {a_\mathcal{A}^2}\mu_i}^{-1} \\
    &\times \brackets{\sum_{i=0}^N \sum_{\ell' m'}\mu_i
    \frac{a_i^{\ell'}}{a_\mathcal{A}^\ell}
    R_{\ell - \ell', m - m'}(\bm r_i)K_{\ell' m'}^{(i)}} \\
  \end{aligned}
\end{equation}
where the unmarked sum limits are $0 \leq l' \leq l$ and $-\ell' \leq m' \leq \ell'$. We also have total mass and center of mass constraints:
\begin{equation}
  \mu_\mathcal{A} = \sum_{i=0}^N \mu_i;  \qquad 0 = \sum_{i=0}^N \mu_i \bm r_i.
\end{equation}
Additional constraints can be imposed on $K_{\ell m}^{(i)}$ if desired. For example, we can require that the lumps be ellipsoids, so that $K_{3 m}^{(i)} = 0$. The most extreme case is spherical lumps, which have $K_{\ell m} = 0$ for $\ell > 0$. $K_{00}=1$ is also guaranteed by definition, meaning that each spherical lump has only five DOF ($a_i$, $\mu_i$, and $\bm r_i$). The substrate has one degree of freedom, since only $\mu_0$ is unknown. Thus, this spherical lumpy model has $5N - 3$ total DOF. Again, the choice of $N$ affects the accuracy and uncertainty of the model results. For the rest of this paper, we use $N=1$ spherical lump for simplicity, corresponding to 2 DOF. In section \ref{sec:density-compare}, we also use $N=2$ spherical lumps which possesses 7 DOF.

@@

  \caption{Cross-sectional slices of the density distributions extracted via the finite element model for the asymmetric (\textit{top two rows}) and symmetric (\textit{bottom two rows}) uniform-density reference asteroids. The finite element model (\textit{first and third rows}) and the lumpy model (\textit{second and fourth rows}) are employed. From left to right, the densities (divided by the average density), deviations from the true density, uncertainties, and significance of the deviations are plotted. These figures are available in animated form in appendix figure \ref{fig:animated-uniform}. Extracted densities are generally within 10\% of the truth.}
  \label{fig:den-uniform}
\end{figure*}

@@

%% Working on cutting down the uncertainties in the large-uncertainty regions

@@

  \caption{Cross-sectional slices of the density distributions extracted via the finite-element (\textit{top}) and lumpy (\textit{bottom}) models for an asteroid with an off-center core. From left to right, the densities, deviations from the true density, uncertainties, and significance of the deviations are plotted. These figures are available in animated form in appendix figure \ref{fig:animated-move}. The lumpy model successfully extracts the core.}
  \label{fig:den-move}
\end{figure*}

@@

  \caption{Cross-sectional slices of the density distributions extracted via the finite-element (\textit{top}) and lumpy (\textit{bottom}) models for an asteroid with a centred core. From left to right, the densities, deviations from the true density, uncertainties, and significance of the deviations are plotted. These figures are available in animated form in appendix figure \ref{fig:animated-sph}. The resulting density distribution is consistent with the density moments but does not represent the true distribution due to degeneracy.}
  \label{fig:den-sph}
\end{figure*}

@@

  \caption{Cross-sectional slices of the density distributions extracted via the finite-element (\textit{top}) and the two-lump lumpy (\textit{bottom}) models for an asteroid with two counterbalancing cores. From left to right, the densities, deviations from the true density, uncertainties, and significance of the deviations are plotted. These figures are available in animated form in appendix figure \ref{fig:animated-double}. The additional lump greatly increases uncertainty, but the resulting distribution is close to accurate.}
  \label{fig:den-double}
\end{figure*}

@@

Figure \ref{fig:scan-oblateness} displays moment uncertainty $\sigma(K_{2 m})$ of the first-order parameters as a function of $\epsilon$ across a reasonable range of central body oblatenesses based on those of Solar System planets \cite{paterLissauer2015}. It also shows linear best-fitting curves for moment uncertainty as a function of oblateness. All other parameters, include $I_\mathcal{B}$ which parametrizes the central body radius, are kept constant. Almost no dependence of $\sigma(K_{\ell m})$ on oblateness $\epsilon$ is apparent, although moment uncertainty does measurably decrease for oblate central bodies.

>>>>

Over the past twenty years, the increase in quantity and quality of sensitive all-sky surveys has prompted the discovery of numerous asteroids. Such advances have been made via ground-based surveys such as the Catalina Sky Survey \citep{larson1998catalina}, Pan-STARRS \citep{kaiser2002pan}, and the Lincoln Near-Earth Asteroid Research project (LINEAR) \citep{stokes2000lincoln}, as well as space-based instruments such as the Wide-field Infrared Survey Explorer (WISE) mission \citep{wright2010wide}. Many of these asteroids are relatively small, but some are kilometre-sized and a few are predicted to closely encounter Earth or other planets in the near future. More encounter candidates are likely to be discovered by new efforts such as the Large-aperture Synoptic Survey Telescope (LSST) \citep{tyson2002large}. Their encounters can then be monitored by global ground-based networks such as the Las Cumbres Observatory (LCO) \citep{brown2013cumbres}. Such ground-based monitoring is typically used to derive the rotation period of an asteroid and its surface properties (see e.g. \cite{10.1093/mnras/stab1252}).

Since the tidal torque acting on an asteroid during an encounter depends on the interior mass distribution, the careful monitoring of angular velocity variations during an encounter also presents a window into the interior properties of asteroids. The gravitational two-body system has been studied in the context of tidal torque to different orders and with several different methods \citep{paul88, SCHEERES2000106, ashenberg07, BOUE2009750, HouMar2017}. Further theoretical studies showed that the tidal torque, observed through angular velocity perturbations, is sensitive to asteroid interior density distribution \citep{Naidu_2015, Makarov2022ChaosOO, scheeres2004evolution}.

Angular velocity perturbations have been observed and used to extract asteroid properties in several cases, including for the 2013 encounter of (367943) Duende with Earth \citep{MOSKOVITZ2020113519, benson2020spin}, and asteroid binaries (3905) Doppler and (617) Patroclus \citep{DESCAMPS2020113726, BERTHIER2020113990}. Orbital and physical properties, including moment of inertia (MOI) ratios have also been extracted for asteroid 99942 Apophis, discovered on June 19, 2004 by R. A. Tucker, D. J. Tholen, and F. Bernardi \citep{giorgini2005recent, giorgini2008predicting, smalley20052004}, and on target to encounter Earth in 2029 \citep{yu2014numerical, hirabayashi2021finite, valvano2022apophis, Lee2022Apophis}. However, density distribution features beyond the MOI ratios have not yet been extracted for any asteroid encounters. More research is needed to study in what cases these effects are observable, and what factors generally inhibit observation of these new features. It seems pivotal to augment previous work on the affect of tidal torque on Apophis' angular velocity in particular \citep{souchay2014rotational, souchay2018changes} so that upcoming observations may constrain these properties and thus improve our predictions.

We address this community need by developing a methodology to translate (1) time series of asteroid angular velocity data into constraints on density moments and (2) constraints on density moments into constraints on an asteroid's density distribution. Other techniques, such as measurement of tidal distortion \citep{RICHARDSON199847}, impact or seismometry experiments \citep{RICHARDSON2005325}, or gradiometry \citep{carroll2018tidal}, may additionally constrain the density distribution. In section \ref{sec:methods}, we introduce the analytical and numerical fundamentals of this methodology. There, we describe a simulation used to integrate the equations of motion and produce synthetic data of angular velocity over time, followed by a Markov Chain Monte Carlo (MCMC) fit process which  extracts density moments from the fit data. We then describe two methods to generate full density distributions from the density moments. In section \ref{sec:results}, we present the results of a series of injection-retrieval tests demonstrating the extent to which the properties of an asteroid chosen to generate synthetic spin data can be retrieved via our methodology. Finally, in section \ref{sec:discussion}, we assess the sensitivity of these constraints to various physical, observational, and methodological parameters to provide guidance for monitoring upcoming close encounters.

@@

\subsection{Extracting density moments from spin data}
\label{sec:fit}
Given synthetic data, an Affine Invariant MCMC Ensemble sampler was used to generate Posterior Probability Distributions (PPDs) from flat priors. We use the Python implementation \texttt{emcee} \citep{foreman2013emcee}. Our parameters were $\gamma_0$, $K_{20}$, $K_{22}$, and $K_{3m}$ (10 in total), bounded by $|\gamma_0| < \pi/4$, and the $K_{2 m}$ bounds given in equation \ref{eqn:parameter-bounds}. Computing $K_{3m}$ for many positive density distributions revealed a $|K_{3m}| \lesssim 0.01$ typically, so we use more conservative bounds on $K_{3 m}$ of $|K_{3 m}| < 1$. In general, spin data is most sensitive to $\gamma_0$ and $K_{2m}$, which we call the ``first-order parameters.'' We call $K_{3m}$ the ``second-order parameters.''

@@

Before the MCMC was run, local minima in the likelihood were found via the Nelder-Mead algorithm implemented in \texttt{scipy} \citep{Gao2012}. Generally, only one local minimum existed, except when $K_{22}=0$ in which case rotational symmetry caused all values of $\gamma_0$ to be degenerate. Ensemble walkers were initialized near the local minimum, distributed by a Gaussian approximation of the likelihood, as determined via the Hessian of the likelihood at the minimum. Due to the high sensitivity of the angular velocity data to density moments, the minimization procedure sometimes failed to isolate the minimum likelihood. Therefore, a simpler simulation without the $K_{3m}$ terms of equation \ref{eqn:tidal-torque} was first used to minimize likelihood as a function of the first-order parameters $\gamma_0$ and $K_{2m}$, and then the full simulation was used to find the second-order parameters $K_{3m}$ with the first-order parameters fixed.

@@

The translation rules of spherical harmonics \citep{Gelderen1998TheSO} give that the density moments in the hybrid frame are
\begin{equation}
  \begin{aligned}
    K_{\ell m} = &\brackets{\sum_{i=0}^N \frac{a_i^2 + r_i^2}
    {a_\mathcal{A}^2}\mu_i}^{-1} \\
    &\times \brackets{\sum_{i=0}^N \sum_{\ell' m'}\mu_i
    \frac{a_i^{\ell'}}{a_\mathcal{A}^\ell}
    R_{\ell - \ell', m - m'}(\bm r_i)K_{\ell' m'}^{(i)}} \\
  \end{aligned}
\end{equation}
where the unmarked sum limits are $0 \leq l' \leq l$ and $-\ell' \leq m' \leq \ell'$. We also have total mass and center of mass constraints:
\begin{equation}
  \mu_\mathcal{A} = \sum_{i=0}^N \mu_i;  \qquad 0 = \sum_{i=0}^N \mu_i \bm r_i.
\end{equation}
Additional constraints can be imposed on $K_{\ell m}^{(i)}$ if desired. For example, we can require that the lumps be ellipsoids, so that $K_{3 m}^{(i)} = 0$. The most extreme case is spherical lumps, which have $K_{\ell m} = 0$ for $\ell > 0$. $K_{00}=1$ is also guaranteed by definition, meaning that each spherical lump has only five DOF ($a_i$, $\mu_i$, and $\bm r_i$). The substrate has one degree of freedom, since only $\mu_0$ is unknown. Thus, this spherical lumpy model has $5N - 3$ total DOF. Again, the choice of $N$ affects the accuracy and uncertainty of the model results. For the rest of this paper, we use $N=1$ spherical lump for simplicity, corresponding to 2 DOF. In section \ref{sec:density-compare}, we also use $N=2$ spherical lumps which possesses 7 DOF.

@@

  \caption{Cross-sectional slices of the density distributions extracted via the finite element model for the asymmetric (\textit{top two rows}) and symmetric (\textit{bottom two rows}) uniform-density reference asteroids. The finite element model (\textit{first and third rows}) and the lumpy model (\textit{second and fourth rows}) are employed. From left to right, the densities (divided by the average density), deviations from the true density, uncertainties, and significance of the deviations are plotted. These figures are available in animated form in the Supplemental Data. Extracted densities are generally within 10\% of the truth.}
  \label{fig:den-uniform}
\end{figure*}

@@

@@

  \caption{Cross-sectional slices of the density distributions extracted via the finite-element (\textit{top}) and lumpy (\textit{bottom}) models for an asteroid with an off-center core. From left to right, the densities, deviations from the true density, uncertainties, and significance of the deviations are plotted. These figures are available in animated form in the Supplemental Data. The lumpy model successfully extracts the core.}
  \label{fig:den-move}
\end{figure*}

@@

  \caption{Cross-sectional slices of the density distributions extracted via the finite-element (\textit{top}) and lumpy (\textit{bottom}) models for an asteroid with a centred core. From left to right, the densities, deviations from the true density, uncertainties, and significance of the deviations are plotted. These figures are available in animated form in the Supplemental Data. The resulting density distribution is consistent with the density moments but does not represent the true distribution due to degeneracy.}
  \label{fig:den-sph}
\end{figure*}

@@

  \caption{Cross-sectional slices of the density distributions extracted via the finite-element (\textit{top}) and the two-lump lumpy (\textit{bottom}) models for an asteroid with two counterbalancing cores. From left to right, the densities, deviations from the true density, uncertainties, and significance of the deviations are plotted. These figures are available in animated form in the Supplemental Data. The additional lump greatly increases uncertainty, but the resulting distribution is close to accurate.}
  \label{fig:den-double}
\end{figure*}

@@

Figure \ref{fig:scan-oblateness} displays moment uncertainty $\sigma(K_{2 m})$ of the first-order parameters as a function of $\epsilon$ across a reasonable range of central body oblatenesses based on those of Solar System planets \citep{paterLissauer2015}. It also shows linear best-fitting curves for moment uncertainty as a function of oblateness. All other parameters, include $I_\mathcal{B}$ which parametrizes the central body radius, are kept constant. Almost no dependence of $\sigma(K_{\ell m})$ on oblateness $\epsilon$ is apparent, although moment uncertainty does measurably decrease for oblate central bodies.
